Odonate abundance changes in response to the creation of a flood reservoir in a former fen

### 2020-12-02 analysis/analysisSS.Rmd
working on results and figures in manuscript

# I want to know how many species there are in a given year, using the nrow function from base
# I will save each as with a corresponding year identifier because I have not yet decided what format I want them all in (dataframe, table, etc.)

@@

>>>>

# I want to know how many species there are in a given year, subsetting using the which function and finding the length of that subset
# I will save each as with a corresponding year identifier because I have not yet decided what format I want them all in (dataframe, table, etc.)

@@

# weird character in genus column name, need to rename it, dplyr function

odonata_Woodwalton <- dplyr::rename(odonata_names, GENUS = ï..GENUS)

### 2020-12-02 manuscript/manuscript.Rmd
working on results and figures in manuscript

---
title: "Odonate abundance changes in response to the creation of a flood reservoir in a former fen"
author: Maggie Vincent
output:
  pdf_document:
    fig_caption: yes

@@

![Map of eastern England with Woodwalton Fen marked.](Woodwalton_Fen_England.jpg)

@@

abunYearSums_1 <- dplyr::rename(abundanceYearSums, year = new_col)
abunYearsSums <- dplyr::rename(abunYearSums_1, total_abundance = rowSums.fullAbundanceDF...c.2..3..4..5..6..7..8..9..10..11...)

par(mfrow=c(1,1))
plot(abunYearsSums$year, abunYearsSums$total_abundance, xlab = "Year", ylab = "Abundance")
abline(lm(abunYearsSums$total_abundance ~ abunYearsSums$year), col="red")

fitTotal <- lm(abunYearsSums$total_abundance ~ abunYearsSums$year)

@@

>>>>

---
title: "Odonate abundance changes in response to the creation of a flood reservoir in a former fen"
author: Maggie Vincent
output:
  pdf_document:
    fig_caption: true

@@

![**Figure 1.** Map of eastern England with Woodwalton Fen marked.](Woodwalton_Fen_England.jpg)

@@

abunYearSums_1 <- dplyr::rename(abundanceYearSums, year = new_col)
abunYearsSums <- dplyr::rename(abunYearSums_1, total_abundance = rowSums.fullAbundanceDF...c.2..3..4..5..6..7..8..9..10..11...)
```

The damselflies and dragonflies in this study are listed in Table 1 for reference. The figure below (Fig. 2), is a linear regression model of the maximum abundance of males in a single day from 1959  to 1988. The abundance corresponds to the sum of the ten species in Table 1 and the p-value for the model was 0.000162.

```{r, echo=FALSE,}

# weird character in genus column name, need to rename it, dplyr function

odonata_Woodwalton <- dplyr::rename(odonata_names, GENUS = ï..GENUS)

knitr::kable(odonata_Woodwalton, caption = 'Table 1. Species and common names of ten Odonates found at Woodwalton Fen between 1959 and 1988.')

```{r, echo=FALSE, fig.align=center, fig.cap='**Figure 2** Linear model of abundance of all ten species from 1959 to 1988. p-value = 0.000162 }
par(mfrow=c(1,1))
plot(abunYearsSums$year, abunYearsSums$total_abundance, xlab = "Year", ylab = "Abundance")
abline(lm(abunYearsSums$total_abundance ~ abunYearsSums$year), col="red")
```

```{r, echo=TRUE}
fitTotal <- lm(abunYearsSums$total_abundance ~ abunYearsSums$year)

@@

```{r, echo=FALSE, fig.cap=paste("Figure 3. Abundances broken down by species with line of best fit.") }
par(mfrow=c(2,5))
plot(bPratenseCompAY$YEAR, bPratenseCompAY$bPratenseAbundance, main = "Brachytron pratense", xlab = "Year", ylab = "Abundance")
abline(lm (bPratenseCompAY$bPratenseAbundance ~ bPratenseCompAY$YEAR), col="red")

plot(pNymphulaCompAY$YEAR, pNymphulaCompAY$pNymphulaAbundance, main = "Pyrrhosa nymphula", xlab = "Year", ylab = "Abundance")
abline(lm (pNymphulaCompAY$pNymphulaAbundance ~ pNymphulaCompAY$YEAR), col="red")

plot(aCyaneaCompAY$YEAR, aCyaneaCompAY$aCyaneaAbundance, main = "Aeshna cyanea", xlab = "Year", ylab = "Abundance")
abline(lm(aCyaneaCompAY$aCyaneaAbundance ~ aCyaneaCompAY$YEAR), col="red")

plot(sSanguineumCompAY$YEAR, sSanguineumCompAY$sSanguineumAbundance, main = "Sympetrum sanguineum", xlab = "Year", ylab = "Abundance")
abline(lm(sSanguineumCompAY$sSanguineumAbundance ~ sSanguineumCompAY$YEAR), col="red")

plot(aGrandisCompAY$YEAR, aGrandisCompAY$aGrandisAbundance, main = "Aeshna grandis", xlab = "Year", ylab = "Abundance")
abline(lm(aGrandisCompAY$aGrandisAbundance ~ aGrandisCompAY$YEAR), col="red")

plot(cPuellaCompAY$YEAR, cPuellaCompAY$cPuellaAbundance, main = "Coenagrion puella", xlab = "Year", ylab = "Abundance")
abline(lm(cPuellaCompAY$cPuellaAbundance ~ cPuellaCompAY$YEAR), col="red")

plot(iElegansCompAY$YEAR, iElegansCompAY$iElegansAbundance, main = "Ischnura elegans", xlab = "Year", ylab = "Abundance")
abline(lm(iElegansCompAY$iElegansAbundance ~ iElegansCompAY$YEAR), col="red")

plot(lSponsaCompAY$YEAR, lSponsaCompAY$lSponsaAbundance, main = "Lestes sponsa", xlab = "Year", ylab = "Abundance")
abline(lm(lSponsaCompAY$lSponsaAbundance ~ lSponsaCompAY$YEAR), col="red")

plot(lQuadrimaculataCompAY$YEAR, lQuadrimaculataCompAY$lQuadrimaculataAbundance, main = "Libellula quadrimaculata", xlab = "Year", ylab = "Abundance")
abline(lm(lQuadrimaculataCompAY$lQuadrimaculataAbundance ~ lQuadrimaculataCompAY$YEAR), col="red")

plot(sStriolatumCompAY$YEAR, sStriolatumCompAY$sStriolatumAbundance, main = "Sympetrum striolatum", xlab = "Year", ylab = "Abundance")
abline(lm(sStriolatumCompAY$sStriolatumAbundance ~ sStriolatumCompAY$YEAR), col="red")

names_p_values <- read.csv("~/GitHub/rrirfinalproject/analysis/names_p_values.csv")

# strange character in column name for species

corrected_names_p_values <- dplyr::rename(names_p_values, Species = ï..SPECIES)

knitr::kable(corrected_names_p_values, caption = 'Table 2. p-values from the LM for each species and all species combined shown in Figure 2. Significance shown with asterisks.')

```{r, echo=FALSE, fig.cap="Figure 4. Number of species in each year of the study period. Note that this is not a comprehensive list of all Odonates at the site, this will be addressed in the discussion."}
# I want to know how many species there are in a given year, subsetting using the which function and finding the length of that subset
# I will save each as with a corresponding year identifier because I have not yet decided what format I want them all in (dataframe, table, etc.)

ASpecies <- length(which(fullAbundanceDF[1,2:11] != 0))
BSpecies <- length(which(fullAbundanceDF[2,2:11] != 0))
CSpecies <- length(which(fullAbundanceDF[3,2:11] != 0))
DSpecies <- length(which(fullAbundanceDF[4,2:11] != 0))
ESpecies <- length(which(fullAbundanceDF[5,2:11] != 0))
FSpecies <- length(which(fullAbundanceDF[6,2:11] != 0))
GSpecies <- length(which(fullAbundanceDF[7,2:11] != 0))
HSpecies <- length(which(fullAbundanceDF[8,2:11] != 0))
ISpecies <- length(which(fullAbundanceDF[9,2:11] != 0))
JSpecies <- length(which(fullAbundanceDF[10,2:11] != 0))
KSpecies <- length(which(fullAbundanceDF[11,2:11] != 0))
LSpecies <- length(which(fullAbundanceDF[12,2:11] != 0))
MSpecies <- length(which(fullAbundanceDF[13,2:11] != 0))
NSpecies <- length(which(fullAbundanceDF[14,2:11] != 0))
OSpecies <- length(which(fullAbundanceDF[15,2:11] != 0))
PSpecies <- length(which(fullAbundanceDF[16,2:11] != 0))
QSpecies <- length(which(fullAbundanceDF[17,2:11] != 0))
RSpecies <- length(which(fullAbundanceDF[18,2:11] != 0))
SSpecies <- length(which(fullAbundanceDF[19,2:11] != 0))
TSpecies <- length(which(fullAbundanceDF[20,2:11] != 0))
USpecies <- length(which(fullAbundanceDF[21,2:11] != 0))
VSpecies <- length(which(fullAbundanceDF[22,2:11] != 0))
XSpecies <- length(which(fullAbundanceDF[23,2:11] != 0))
YSpecies <- length(which(fullAbundanceDF[24,2:11] != 0))
ZSpecies <- length(which(fullAbundanceDF[25,2:11] != 0))
AASpecies <- length(which(fullAbundanceDF[26,2:11] != 0))
ABSpecies <- length(which(fullAbundanceDF[27,2:11] != 0))
ACSpecies <- length(which(fullAbundanceDF[28,2:11] != 0))
ADSpecies <- length(which(fullAbundanceDF[29,2:11] != 0))
AESpecies <- length(which(fullAbundanceDF[30,2:11] != 0))

# number of species in chronological order
# I'll reuse the years vector I created earlier to make a data.frame

speciesChrono <- c(ASpecies, BSpecies, CSpecies, DSpecies, ESpecies, FSpecies, GSpecies, HSpecies, ISpecies, JSpecies, KSpecies, LSpecies, MSpecies, NSpecies, OSpecies, PSpecies, QSpecies, RSpecies, SSpecies, TSpecies, USpecies, VSpecies, XSpecies, YSpecies, ZSpecies, AASpecies, ABSpecies, ACSpecies, ADSpecies, AESpecies)

speciesByYear <- data.frame(YEAR, speciesChrono)

# look at it graphically

par(mfrow=c(1,1))
plot(speciesByYear$YEAR, speciesByYear$speciesChrono, xlab = "Year", ylab = "Species Number")
